---
title: "Lab Report 1"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
```

```{r}
knitr::opts_chunk$set(
        echo = TRUE,
        warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() +theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r}
# importing and cleaning the data
weekly_ili = read_csv(file = "./data/da_ILINet.csv") |> 
  janitor::clean_names() |> 
  select(-florida)
```

# Question 1
Explore the dataset and answer the following questions:

## 1.1
What variables does the dataset contain? How many weeks are included for each variable in the dataset? (Hint: what function gives the column names? What function gives the number of rows in a data frame?)

The data set contain a date variable to denote one day per week from September 30, 2018 to September 22, 2019, for a total of 52 weeks. The other variables are 50 U.S. states (although there is no data recorded for  Florida) and U.S. territories such as Puerto Rico and the Virgin Islands. The dataset also includes two city variables: Washington D.C. and New York City.

## 1.2
What state are you from? What was the ILI in the 2nd week of the season for your state? (Note if you are from out of the States or Florida (the CDC did not have data for Florida that season), please use New York City instead).

I am from New York, and the ILI in the 2nd week of the season was 982.

## 1.3
Suppose you store the dataset in a variable called 'ili'. What do the outputs from the following commands represent? NOTE: Specify the date(s) and location(s); do not copy & paste the output.

```{r}
ili = weekly_ili |> 
  pivot_longer(
    cols = alabama:wyoming,
    names_to = "state",
    values_to = "ili"
  )
```

The outputs from the following commands represent the influenza-like illness per 100,000 cases recorded for that week.

### part a
ili [2, 6]

### part b
ili [30, 5:14]

### part c
ili [, 5]

### part d
ili [36, 'New.York.City'] (Note: R does not 'like' space in variable names, so replace space with '.' here)

## 1.4
What are the mean and maximum ILI levels for New York City, respectively? (hint: you can use the 'summary' function to print the summary statistics for each column. Or use the function 'mean' and 'max' to get the results respectively)

The mean ILI level for New York City is 2290.
The maximum ILI level for New York City is 4383.

## 1.5
When did New York City have its peak ILI (i.e. maximum ILI) that flu season?

New York City had its peak ILI on December 23, 2018 that flu season.


# Question 2
Plot the weekly ILI in New York City (column: New.York.City) over time.

```{r}
nyc = ggplot(weekly_ili, aes(x = date, y = new_york_city, color = date)) +
  geom_point(alpha = 0.6) +
  geom_smooth() +
  labs(title = "Weekly ILI in New York City", 
       x = "Date", 
       y = "Influenza-like Illness (per 100,000)")
```


# Question 3
New York City is in New York State. To compare ILI in NYC to that in NY State, superimpose the weekly ILI for NY State to the above plot (Q2).

```{r}
ny_state = ggplot(weekly_ili, aes(x = date, y = new_york, color = date)) +
  geom_point(alpha = 0.6) +
  geom_smooth() +
  labs(title = "Weekly ILI in New York State", 
       x = "Date", 
       y = "Influenza-like Illness (per 100,000)")
```

```{r}
nyc + ny_state
```

# Question 4
When you have multiple variables/categories and each has a lot of data points, boxplot is an effective way to display the distributions of the variables/categories. Use boxplot to show the variation in ILI among all states/cities for each week in the last flu season. 

```{r}
ggplot(ili, aes(x = date, y = ili, color = state)) +
  geom_boxplot()
```


# Math Operations & Function
# Question 5
We know that the solutions of the quadratic equation ax^2 + bx + c = 0 are:
Write a script in R to find the solutions for the following equations:


## part a
2x^2 + 10x + 3 = 0

```{r}
a = 2
b = 10
c = 3
```

```{r}
discriminant = b^2 - 4* a * c
```

```{r}
if (discriminant > 0) {
  root1 = (-b + sqrt(discriminant)) / (2 * a)
  root2 = (-b - sqrt(discriminant)) / (2 * a)
  print(c(root1, root2))
} else if (discriminant == 0) {
  root = -b / (2 * a)
  print(root)
} else {
  realPart = -b / (2 * a)
  imaginaryPart = sqrt(-discriminant) / (2 * a)
  print(paste(realPart, "+", imaginaryPart, "i"))
  print(paste(realPart, "-", imaginaryPart, "i"))
}
```


## part b
5x^2 - 6x + 1 = 0

```{r}
a = 5
b = -6
c = 1
```

```{r}
discriminant = b^2 - 4* a * c
```

```{r}
if (discriminant > 0) {
  root1 = (-b + sqrt(discriminant)) / (2 * a)
  root2 = (-b - sqrt(discriminant)) / (2 * a)
  print(c(root1, root2))
} else if (discriminant == 0) {
  root = -b / (2 * a)
  print(root)
} else {
  realPart = -b / (2 * a)
  imaginaryPart = sqrt(-discriminant) / (2 * a)
  print(paste(realPart, "+", imaginaryPart, "i"))
  print(paste(realPart, "-", imaginaryPart, "i"))
}
```


# Question 6
To make implementation of Question 5 easier, create a function to find the solutions for any quadratic equation with constants a, b, and c. Your function should be able to return the two solutions once you specify the three constants, For instance, suppose you name the function Fn_sol_quadratic, the command Fn_sol_quadratic(a=2, b=10, c=3) will return the two solutions for equation 2x^2 +10x +3 = 0

```{r}
quadratic = function (a, b, c) {
  
  discriminant = b^2 - 4*a*c
  
  if (discriminant > 0) {
    root1 = (-b + sqrt(discriminant)) / (2*a)
    root2 = (-b - sqrt(discriminant)) / (2*a)
    return(c(root1, root2))
    
  } else if (discriminant == 0) {
    
    root = -b / (2*a)
    return(root)
    
    } else {
      real_part = -b / (2*a)
      imaginary_part = sqrt(-discriminant) / (2*a)
      return(c(real = real_part, imaginary = imaginary_part))
  }
}
```

2x^2 + 10x + 3 = 0
```{r}
quadratic(2, 10, 3)
```

5x^2 - 6x + 1 = 0
```{r}
quadratic(5, -6, 1)
```

